---
title: "In-class Exercise 6"
author: "Jin Jiaying"
date: "03 Oct 2026"
date-modified: last-modified
---

# Loading the necessary R packages
```{r}
pacman::p_load(sf, sfdep, tmap, plotly, tidyverse, Kendall)
```

# Getting the Data Into R Environment
```{r}
hunan <- st_read(
  dsn = "in-class_ex6_data/geospatial", 
  layer = "Hunan") %>% 
  st_transform(crs = 32650)
```
```{r}
GDPPC <- read_csv(
  "in-class_ex6_data/aspatial/Hunan_GDPPC.csv")
```
## Create a Time Series Cube
```{r}
GDPPC_st <- spacetime(GDPPC, #data in data frame format
                      hunan, # geometry
                      .loc_col = "County", #location identifier
                      .time_col = "Year" #time column
                      ) 

is_spacetime_cube(GDPPC_st)
```
:::{.callout-note title="space()"}
Make sure **time column** should be involved in **data which is in data frame format**.
**Location identifiers** should be involved in **geometry**.
After that, it is always good to test if it is a cube using **is_spacetime_cube()**.
:::

# Computing Gi*
```{r}
GDPPC_nb <- GDPPC_st %>%
  activate("geometry") %>%
  mutate(nb = include_self(
    st_contiguity(geometry)),
    wt = st_inverse_distance(nb, 
                             geometry, 
                             scale = 1,
                             alpha = 1),
    .before = 1) %>%
  set_nbs("nb") %>%
  set_wts("wt")

gi_stars <- GDPPC_nb %>% 
  group_by(Year) %>% 
  mutate(gi_star = local_gstar_perm(
    GDPPC, nb, wt)) %>% 
  tidyr::unnest(gi_star)
```

# Mann-Kendall Test
```{r}
cbg <- gi_stars %>% 
  ungroup() %>% 
  filter(County == "Changsha") |> 
  select(County, Year, gi_star)

ggplot(data = cbg, 
       aes(x = Year, 
           y = gi_star)) +
  geom_line() +
  theme_light()
```

```{r}
p <- ggplot(data = cbg, 
       aes(x = Year, 
           y = gi_star)) +
  geom_line() +
  theme_light()

ggplotly(p)
```

```{r}
cbg %>%
  summarise(mk = list(
    unclass(
      Kendall::MannKendall(gi_star)))) %>% 
  tidyr::unnest_wider(mk)
```

```{r}
ehsa <- gi_stars %>%
  group_by(County) %>%
  summarise(mk = list(
    unclass(
      Kendall::MannKendall(gi_star)))) %>%
  tidyr::unnest_wider(mk)
head(ehsa)

emerging <- ehsa %>% 
  arrange(sl, abs(tau)) %>% 
  slice(1:10)
head(emerging)
```

# Performing Emerging Hotspot Analysis (EHSA)
```{r}
ehsa <- emerging_hotspot_analysis(
  x = GDPPC_st, 
  .var = "GDPPC", 
  k = 1, 
  nsim = 99
)
```

## Visualising the distribution of EHSA classes
```{r}
ggplot(data = ehsa,
       aes(x = classification)) +
  geom_bar()
```

## Visualising EHSA
```{r}
hunan_ehsa <- hunan %>%
  left_join(ehsa,
            by = join_by(County == location))

ehsa_sig <- hunan_ehsa  %>%
  filter(p_value < 0.05)
tmap_mode("plot")
tm_shape(hunan_ehsa) +
  tm_polygons() +
  tm_borders(fill_alpha = 0.5) +
tm_shape(ehsa_sig) +
  tm_fill("classification") + 
  tm_borders(fill_alpha = 0.4)
```
