---
title: "In-class Exercise 4"
author: "Jin Jiaying"
date: "19 Sep 2026"
date-modified: last-modified
---

# Loading the necessary R packages

```{r}
pacman::p_load(sf, ggstatsplot, spdep, tmap, tidyverse, knitr, GWmodel)
```

# Importing Dataset

```{r}
hunan_sf <- st_read(dsn = "in-class_ex4_data", layer = "Hunan")

hunan2012 <- read_csv("in-class_ex4_data/Hunan_2012.csv")

hunan <- left_join(hunan_sf, hunan2012)%>%
  select(1:3, 7, 15, 16, 31, 32)

hunan_alt <- left_join(hunan_sf, hunan2012)%>%
  select("NAME_2" , "ID_3", "NAME_3", "County", "GDPPC", "GIO", "Agri", "Service", "geometry")
```

::: {.callout-note title="Joining Technics"}
A general way to joining data table is using **by = join_by()**, to identify which variable used for join.
:::

# Mapping GDPPC

```{r}
basemap <- tm_shape(hunan) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan, "GDPPC") +
  tm_layout(
    legend.position = c("left", "bottom"))
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```

::: {.callout-note title="About qtm()"}
qtm() function has a default colour design as blue. Make sure to adjust the legend position, to not blocking the map.
:::

# Converting to Spatial Polygon Dataframe

```{r}
hunan_sp <- hunan%>%
  as_Spatial()
```

# Geographically Weighted Summary Statistics (GSWW) with adaptive bandwidth

```{r}
bw_CV <- bw.gwr(GDPPC ~ 1,
                data = hunan_sp,
                approach = "CV",
                adaptive = TRUE,
                kernel = "bisquare",
                longlat = T)

bw_CV
```

```{r}
bw_AIC <- bw.gwr(GDPPC ~ 1,
                data = hunan_sp,
                approach = "AICc",
                adaptive = TRUE,
                kernel = "bisquare",
                longlat = T)

bw_AIC
```

::: {.callout-note title="Bandwidth selection for basic GWR"}
bw.gwr() is a function for automatic bandwidth selection to calibrate a basic GWR model.

Usage: bw.gwr(formula, data, approach="CV", kernel="bisquare", adaptive=FALSE, p=2, theta=0, longlat=F, dMat, parallel.method=F,parallel.arg=NULL) \*CV here stands for Cross-Verification

CV vs AIC:

- AICc: Select a bandwidth by balancing goodness-of-fit and model complexity. A lower AICc indicates a better trade-off and helps reduce overfitting.

- Cross-Validation (CV): Select a bandwidth by minimising out-of-sample prediction error, commonly using leave-one-out prediction. A lower CV score indicates better predictive performance.

In summary

| Objective                       | More Suitable |
|:--------------------------------|:-------------:|
| Interpret spatial relationships |     AICc      |
| Model selection                 |     AICc      |
| Focus on predictive performance |      CV       |
:::

# Computing geographically weighted summary statistics

```{r}
gwstat <- gwss(data = hunan_sp,
               vars = "GDPPC",
               bw = bw_AIC,
               kernel = "bisquare",
               quantile = TRUE,
               adaptive = TRUE,
               longlat = T)
```

# Preparing the output data

```{r}
gwstat_df <- as.data.frame(gwstat$SDF)
```

```{r}
hunan_gstat <- cbind(hunan, gwstat_df)
```

\::: {.callout-note title: Why using **cbind()** here?} The reason cannot using leftjoin as previous is there is **no unique identifier** in gwstat_df dataset. Thus cbind() is used, by combining using same order. (Sequence never change)

Thus, if sorting is done, cannot use cbind() in this case, since sequence has been changed. \:::

# Visualising geographically weighted summary statistics

```{r}
tm_shape(hunan_gstat) +
  tm_polygons(fill = "GDPPC_LM",
              fill.scale = tm_scale_intervals(
                style = "quantile",
                n = 5,
                values = "brewer.greens"),
              fill.legend = tm_legend(
                title = "Local Mean")) + 
  tm_borders(fill_alpha = 0.5) + 
  tm_title("Distribution of geographically weighted mean") + 
  tm_layout(legend.position = c("left", "bottom"))
```
