---
title: "Hands-on Exercise 4"
author: "Jin Jiaying"
date: "12 Sep 2026"
date-modified: last-modified
---

# Loading the necessary R packages

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse, knitr)
```

::: {.callout-note title="Note on spdep package"}
**spdep package** consists functions including poly2nb(), knearneigh(), nb2listw() and lag.listw(). It is mainly for neighbour, spatial weights and spatial lag built-up.
:::

# 4 Spatial Weights and Applications

## 4.1 Getting the Data Into R Environment

```{r}
hunan <- st_read(dsn = "hands-on_ex4_data", layer = "Hunan")

hunan2012 <- read_csv("hands-on_ex4_data/Hunan_2012.csv")

hunan <- left_join(hunan,hunan2012) %>%
  select(1:4, 7, 15)
```

## 4.2 Visualising Regional Development Indicator

::: {.callout-note title="Note on Spatial Weight Matrix"}
Spatial Weight Matrix is transformation from geographical relationship into mathematical relationship.
:::

```{r}
basemap <- tm_shape(hunan) +
  tm_polygons() +
  tm_text("NAME_3", size=0.5)

gdppc <- qtm(hunan, "GDPPC") +
  tm_layout(
    legend.position = c("left", "bottom"))
tmap_arrange(basemap, gdppc, asp=1, ncol=2)
```

## 4.3 Computing Contiguity Spatial Weights

```{r}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)

wm_q[[1]]

hunan$County[1]

hunan$NAME_3[c(2,3,4,57,85)]

nb1 <- wm_q[[1]]
nb1 <- hunan$GDPPC[nb1]
nb1

str(wm_q)
```

::: {.callout-note title="Why to check neighbour"}
Checking neighbours helps validate the spatial weights before using them in further spatial analysis.
:::

```{r}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

::: {.callout-note title="Queen vs Rook Contiguity"}
Both methods define neighbours based on whether polygons share boundaries, but they use different rules.

**Queen**

- Neighbour Definition: Share an **edge or a vertex**

- Main Advantage: More inclusive and usually gives better connectivity Main

- Main Limitation: Corner-touching areas may not have meaningful interaction

**Rook**

- Neighbour Definition: Share an **edge only**

- Main Advantage: Uses a stricter and more direct boundary relationship

- Main Limitation: May exclude some nearby areas that only meet at a corner

**How to Choose?**

- Use Queen contiguity when general spatial proximity and connectivity are important.

- Use Rook contiguity when interaction is expected mainly through a shared boundary.

- The choice should depend on the spatial process and research question, not on which method produces more neighbours.

To summarize, Queen is more inclusive, while Rook is more restrictive.
:::

```{r}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])

latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])

coords <- cbind(longitude, latitude)

head(coords)
```

### 4.3.1 Plotting Queen contiguity based neighbours map

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```

### 4.3.2 Plotting Rook contiguity based neighbours map

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

### 4.3.3 Plotting both Queen and Rook contiguity based neighbours maps

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey", main="Queen Contiguity")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
plot(hunan$geometry, border="lightgrey", main="Rook Contiguity")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## 4.4 Computing distance based neighbours

```{r}
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)

wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62

str(wm_d62)

table(hunan$County, card(wm_d62))

n_comp <- n.comp.nb(wm_d62)
n_comp$nc

table(n_comp$comp.id)
```

```{r}
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE)
plot(k1, coords, add=TRUE, col="red", length=0.08)
```

```{r}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey", main="1st nearest neighbours")
plot(k1, coords, add=TRUE, col="red", length=0.08)
plot(hunan$geometry, border="lightgrey", main="Distance link")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6)
```

```{r}
knn6 <- knn2nb(knearneigh(coords, k=6))
knn6

str(knn6)
```

```{r}
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## 4.5 Weights based on IDW

```{r}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
ids
```

## 4.6 Row-standardised Weights Matrix

```{r}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q

rswm_q$weights[10]

rswm_ids <- nb2listw(wm_q, glist=ids, style="B", zero.policy=TRUE)
rswm_ids

rswm_ids$weights[1]

summary(unlist(rswm_ids$weights))
```

## 4.7 Application of Spatial Weight Matrix

```{r}
GDPPC.lag <- lag.listw(rswm_q, hunan$GDPPC)
GDPPC.lag

nb1 <- wm_q[[1]]
nb1 <- hunan$GDPPC[nb1]
nb1

lag.list <- list(hunan$NAME_3, lag.listw(rswm_q, hunan$GDPPC))
lag.res <- as.data.frame(lag.list)
colnames(lag.res) <- c("NAME_3", "lag GDPPC")
hunan <- left_join(hunan,lag.res)

head(hunan)
```

```{r}
gdppc <- qtm(hunan, "GDPPC")
lag_gdppc <- qtm(hunan, "lag GDPPC")
tmap_arrange(gdppc, lag_gdppc, asp=1, ncol=2)
```

```{r}
b_weights <- lapply(wm_q, function(x) 0*x + 1)
b_weights2 <- nb2listw(wm_q, 
                       glist = b_weights, 
                       style = "B")
b_weights2

lag_sum <- list(hunan$NAME_3, lag.listw(b_weights2, hunan$GDPPC))
lag.res <- as.data.frame(lag_sum)
colnames(lag.res) <- c("NAME_3", "lag_sum GDPPC")

lag_sum
```

```{r}
hunan <- left_join(hunan, lag.res)

gdppc <- qtm(hunan, "GDPPC")
lag_sum_gdppc <- qtm(hunan, "lag_sum GDPPC")
tmap_arrange(gdppc, lag_sum_gdppc, asp=1, ncol=2)
```

```{r}
wm_qs <- include.self(wm_q)

wm_qs[[1]]

wm_qs <- nb2listw(wm_qs)
wm_qs

lag_w_avg_gpdpc <- lag.listw(wm_qs, hunan$GDPPC)
lag_w_avg_gpdpc

lag.list.wm_qs <- list(hunan$NAME_3, lag.listw(wm_qs, hunan$GDPPC))
lag_wm_qs.res <- as.data.frame(lag.list.wm_qs)
colnames(lag_wm_qs.res) <- c("NAME_3", "lag_window_avg GDPPC")

hunan <- left_join(hunan, lag_wm_qs.res)

hunan %>%
  select("County", 
         "lag GDPPC", 
         "lag_window_avg GDPPC") %>%
  kable()

w_avg_gdppc <- qtm(hunan, "lag_window_avg GDPPC")
tmap_arrange(lag_gdppc, w_avg_gdppc, asp=1, ncol=2)
```

```{r}
wm_qs <- include.self(wm_q)
wm_qs

b_weights <- lapply(wm_qs, function(x) 0*x + 1)
b_weights[1]

b_weights2 <- nb2listw(wm_qs, 
                       glist = b_weights, 
                       style = "B")
b_weights2

w_sum_gdppc <- list(hunan$NAME_3, lag.listw(b_weights2, hunan$GDPPC))
w_sum_gdppc

w_sum_gdppc.res <- as.data.frame(w_sum_gdppc)
colnames(w_sum_gdppc.res) <- c("NAME_3", "w_sum GDPPC")

hunan <- left_join(hunan, w_sum_gdppc.res)

hunan %>%
  select("County", "lag_sum GDPPC", "w_sum GDPPC") %>%
  kable()

w_sum_gdppc <- qtm(hunan, "w_sum GDPPC")
tmap_arrange(lag_sum_gdppc, w_sum_gdppc, asp=1, ncol=2)
```

::: {.callout-note title="Spatial Lagged Variable vs Spatial Window Sum"}
**Spatial Laggged Variable** exclude focal area itself, while **Spatial Window Sum** include focal area itself.

**Sptial Lagged Variable** is suitable for competitive benchmarking & focal area's own performance.

**Spatial Window Sum** is suitable for additive variables.

In general, **Spatial Window Sum** helps identify where resources are needed based on current local demand, while **Spatial Lagged Variable** helps identify areas that may require strategic attention based on surrounding conditions.
:::
