---
title: "Hands-on Exercise 1"
Author: Jin Jiaying
Date: 28 Aug 2026
---


## Loading the necessary R packages

```{r}
pacman::p_load(sf, tmap, tidyverse, rvest)
```

# 1 Geospatial Data Science with R

## 1.1 Importing Geospatial Data

```{r}
mpsz <- st_read("hands-on_ex1_data/MasterPlan2014SubzoneBoundaryNoSea.geojson")

cyclingpath <- st_read("hands-on_ex1_data/CyclingPathNetworkGEOJSON.geojson")

preschool <- st_read("hands-on_ex1_data/PreSchoolsLocation.geojson")
```

## 1.2 Checking the Content of A Simple Feature Data Frame

```{r}
st_geometry(mpsz)

glimpse(mpsz)

head(mpsz, n=5)
```

## 1.3 Plotting the Geospatial Data

```{r}
plot(mpsz)
```
```{r}
plot(st_geometry(mpsz))
```
```{r}
plot(mpsz["PLN_AREA_N"])
```
```{r}
plot(st_geometry(mpsz))
plot(st_geometry(preschool), add = TRUE)
```

## 1.4  Working with Projection

```{r}
st_crs(mpsz)
mpsz <- st_transform(mpsz, crs = 3414)
st_crs(mpsz)

preschool <- st_transform(preschool, crs = 3414)
st_crs(preschool)
```
```{r}
plot(st_geometry(mpsz))
plot(st_geometry(preschool), add = TRUE)
```

## 1.5 Importing and Converting An Aspatial Data

```{r}
listings <- read_csv("hands-on_ex1_data/listings.csv")
list(listings)
```

```{r}
listings_sf <- st_as_sf(listings, coords = c("longitude", "latitude"), crs=4326) %>%
  st_transform(crs = 3414)
glimpse(listings_sf)
```
```{r}
plot(st_geometry(mpsz))
plot(st_geometry(listings_sf), add = TRUE)
```


## 1.6  Geoprocessing with sf package

### 1.6.1 Use case 1: Land acquisition analysis

```{r}
buffer_cycling <- st_buffer(cyclingpath, dist = 5, nQuadSegs = 30)
buffer_cycling <- buffer_cycling %>%
  mutate(
    AREA = st_area(st_geometry(buffer_cycling))
  )
sum(buffer_cycling$AREA)
```
```{r}
mpsz_selected <- mpsz %>%
  filter(SUBZONE_N == "TAMPINES WEST")
buffer_cycling <- st_transform(buffer_cycling, crs = 3414)
mpsz_selected <- st_transform(mpsz_selected, crs = 3414)

buffer_cycling_selected <- st_intersection(buffer_cycling, mpsz_selected)
plot(st_geometry(mpsz_selected))
plot(st_geometry(buffer_cycling_selected), add = TRUE)
```

### 1.6.2 Use case 2: To determine the number of pre-schools by planning subzone

```{r}
mpsz$`PreSch Count`<- lengths(st_intersects(mpsz, preschool))
summary(mpsz$`PreSch Count`)
top_n(mpsz, 1, `PreSch Count`)
```
```{r}
mpsz$Area <- mpsz %>%
  st_area()
mpsz <- mpsz %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```
```{r}
hist(mpsz$`PreSch Density`)
```
```{r}
ggplot(data=mpsz, aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, color="black", fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?", subtitle= "There are many planning subzones with a single pre-school, on the other hand, \nthere are seven planning subzones with at least 30 or more pre-schools", x = "Pre-school density (per km sq)", y = "Frequency")
```
```{r}
ggplot(data=mpsz, aes(y = `PreSch Count`, x= as.numeric(`PreSch Density`)))+
  geom_point(color="black", fill="light blue") + xlim(0, 40) + ylim(0, 40) +
  labs(title = "", x = "Pre-school density (per km sq)", y = "Pre-school count")
```

# 2 Thematic Mapping and GeoVisualisation with R

## 2.1 Importing Data into R

```{r}
mpsz2 <- st_read("hands-on_ex1_data/MasterPlan2019SubzoneBoundaryNoSeaGEOJSON.geojson") %>%
  st_transform(crs = 3414)

mpsz2 <- mpsz2 %>%
  select(REGION_N, PLN_AREA_N, SUBZONE_N, SUBZONE_C, geometry)
```
```{r}
mpsz2
```
```{r}
popdata <- read_csv("hands-on_ex1_data/respopagesextod2024.csv")

popdata2024 <- popdata %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup()%>%
  pivot_wider(names_from=AG, 
              values_from=POP) %>%
  mutate(YOUNG = rowSums(.[3:6])
         +rowSums(.[12])) %>%
mutate(`ECONOMY ACTIVE` = rowSums(.[7:11])+
rowSums(.[13:15]))%>%
mutate(`AGED`=rowSums(.[16:21])) %>%
mutate(`TOTAL`=rowSums(.[3:21])) %>%  
mutate(`DEPENDENCY` = (`YOUNG` + `AGED`)
/`ECONOMY ACTIVE`) %>%
  select(`PA`, `SZ`, `YOUNG`, 
       `ECONOMY ACTIVE`, `AGED`, 
       `TOTAL`, `DEPENDENCY`)

popdata2024 <- popdata2024 %>%
  mutate_at(.vars = vars(PA, SZ), 
          .funs = list(toupper)) %>%
  filter(`ECONOMY ACTIVE` > 0)

mpsz_pop2024 <- left_join(mpsz2, popdata2024,
                          by = c("SUBZONE_N" = "SZ"))

write_rds(mpsz_pop2024, "hands-on_ex1_data/mpsz_pop2024.rds")
```

## 2.2 Choropleth Mapping Geospatial Data Using tmap

```{r}
tmap_mode("plot")
qtm(shp = mpsz_pop2024, fill = "DEPENDENCY")
```
```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY",
              fill.scale = tm_scale_intervals(
                style = "quantile", 
                n = 5,
                values = "brewer.blues"),
              fill.legend = tm_legend(
                title = "Dependency ratio")) +
  tm_title("Distribution of Dependency Ratio by planning subzone") +
  tm_layout(frame = TRUE) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and Population data from Department of Statistics DOS", 
             position = c("left", "bottom"))
```
```{r}
tm_shape(mpsz_pop2024) + tm_polygons()

tm_shape(mpsz_pop2024)+ tm_polygons(fill = "DEPENDENCY")

tm_shape(mpsz_pop2024)+ tm_fill("DEPENDENCY")

tm_shape(mpsz_pop2024)+ tm_fill("DEPENDENCY") + tm_borders()

tm_shape(mpsz_pop2024)+
  tm_fill("DEPENDENCY") +
  tm_borders(col = "grey60",
             lwd = 0.1,
             lty = "dashed")
```
```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", n = 5)) +
  tm_borders(fill_alpha = 0.5)

tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "equal", n = 5)) +
  tm_borders(fill_alpha = 0.5)
```
### 2.2.1 Different classification methods

```{r}
map_quantile <- tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", n = 5, values = "brewer.blues"),
    fill.legend = tm_legend(title = "Dependency Ratio")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Quantile")
map_quantile

map_equal <- tm_shape(mpsz_pop2024) +
  tm_polygons(
    fill = "DEPENDENCY",
    fill.scale = tm_scale_intervals(
      style = "equal",
      n = 5,
      values = "brewer.blues"
    ),
    fill.legend = tm_legend(
      title = "Dependency Ratio"
    )
  ) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Equal Interval")
map_equal

map_pretty <- tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "pretty", n = 5, values = "brewer.blues"),
    fill.legend = tm_legend(title = "Dependency Ratio")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Pretty")
map_pretty

map_jenks <- tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "jenks", n = 5, values = "brewer.blues"),
    fill.legend = tm_legend(title = "Dependency Ratio")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Jenks")
map_jenks
```

### 2.2.2 Similar classification method (quantile) but with different numbers of classes

```{r}
map_q2 <- tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", n = 2, values = "brewer.blues")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Quantile: 2 Classes")
map_q2

map_q6 <- tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", n = 6, values = "brewer.blues")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Quantile: 6 Classes")
map_q6

map_q10 <- tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", n = 10, values = "brewer.blues")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Quantile: 10 Classes")
map_q10

map_q20 <- tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", n = 20, values = "brewer.blues")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Quantile: 20 Classes")
map_q20
```

### 2.2.3 Plotting choropleth map with custome break

```{r}
summary(mpsz_pop2024$DEPENDENCY)

tm_shape(mpsz_pop2024)+
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(breaks = c(0, 0.60, 0.70, 0.80, 0.90, 1.00))) +
  tm_borders(fill_alpha = 0.5)
```
### 2.2.4 Colour Scheme

```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", n = 5, values = "brewer.greens")) +
  tm_borders(fill_alpha = 0.5)

tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", n = 5, values = "-brewer.greens")) +
  tm_borders(fill_alpha = 0.5)

tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", n = 5)) +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: data.gov.sg & singstat", position = c("left", "bottom"))
```

### 2.2.5 Map Layout

```{r}
tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", n = 5), fill.legend = tm_legend(title = "Dependency ratio")) +
  tm_pos_auto_in() +
  tm_borders(fill_alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scalebar() +
  tm_grid(lwd = 0.1, alpha = 0.2) +
  tm_credits("Source: data.gov.sg & singstat", position = c("left", "bottom"))

tm_shape(mpsz_pop2024) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", n = 5, values = "-brewer.greens")) + 
  tm_borders(fill_alpha = 0.5) + 
  tmap_style("natural")
```
```{r}
tmap_style("white")
```

## 2.3 Drawing Small Multiple Choropleth Maps

```{r}
tm_shape(mpsz_pop2024) + 
  tm_polygons(fill = c("YOUNG", "AGED"), fill.legend = tm_legend(position = tm_pos_in("right", "bottom")), fill.scale = tm_scale_intervals(style = "equal",  n = 5, values = "brewer.blues")) +
  tm_borders(fill_alpha = 0.5) +
  tmap_style("natural")

youngmap <- tm_shape(mpsz_pop2024)+ 
  tm_polygons(fill = "YOUNG", fill.legend = tm_legend(position = tm_pos_in("right", "bottom"), item.height = 0.8), fill.scale = tm_scale_intervals(style = "quantile", values = "brewer.blues")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Distribution of young population")
                
agedmap <- tm_shape(mpsz_pop2024)+ 
  tm_polygons(fill = "AGED", fill.legend = tm_legend(position = tm_pos_in("right", "bottom"), item.height = 0.8), fill.scale = tm_scale_intervals(style = "quantile", values = "brewer.blues")) +
  tm_borders(fill_alpha = 0.5) +
  tm_title("Distribution of aged population")

tmap_arrange(youngmap, agedmap, asp=1, ncol=2)

tm_shape(mpsz_pop2024) +
  tm_fill(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", values = "brewer.blues")) + 
  tm_facets(by = "REGION_N", nrow = 2, ncols = 3, free.coords=TRUE, drop.units=TRUE) +
  tm_layout(legend.show = TRUE, title.position = c("center", "center"), title.size = 20) +
  tm_borders(fill_alpha = 0.5)
```

## 2.4 Mappping Spatial Object Meeting a Selection Criterion

```{r}
mpsz_pop2024 %>%
  filter(REGION_N == "CENTRAL REGION") %>%
  tm_shape() +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", values = "brewer.greens"), fill.legend = tm_legend()) +
  tm_borders(fill_alpha = 0.5)
```

## 2.5 Complementing Thematic Map with Statistical Chart

```{r}
mpsz_pop2024 %>%
  filter(REGION_N == "CENTRAL REGION") %>%
  tm_shape() +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", values = "brewer.greens"), fill.legend = tm_legend(), fill.chart = tm_chart_box()) +
  tm_borders() +
  tm_layout(asp = 0.8)

mpsz_selected <- mpsz_pop2024 %>%
  filter(REGION_N == "CENTRAL REGION")
stats <- boxplot.stats(mpsz_selected$DEPENDENCY)
outlier_vals <- stats$out
outlier_sf <- mpsz_selected[mpsz_selected$DEPENDENCY %in% outlier_vals, ]
tm_shape(mpsz_selected) +
  tm_polygons(fill = "DEPENDENCY", fill.scale = tm_scale_intervals(style = "quantile", values = "brewer.blues"), fill.legend = tm_legend(), fill.chart = tm_chart_box()) +
  tm_borders(fill_alpha = 0.5) +
tm_shape(outlier_sf) +
  tm_borders(col = "red", lwd = 2) +
  tm_text("SUBZONE_N", col = "red", size = 0.7) +
  tm_layout(asp = 0.8)
```


## 2.6 Creating Interactive Map

```{r}
region_selected <- mpsz_pop2024 %>%
  filter(REGION_N == "CENTRAL REGION")
region_bbox <- st_bbox(region_selected)

stats <- boxplot.stats(region_selected$DEPENDENCY)
outlier_vals <- stats$out
outlier_sf <- region_selected[region_selected$DEPENDENCY %in% outlier_vals, ]

tmap_mode("view")
tm_shape(region_selected, bbox = region_bbox) +
  tm_fill("DEPENDENCY", id = "SUBZONE_N", popup.vars = c("Name" = "SUBZONE_N", "Dependency" = "DEPENDENCY")) +
  tm_borders() +
  tm_shape(outlier_sf) +
  tm_borders(col = "red", lwd = 2)

tmap_mode("plot")
```

```{r}
region_selected <- mpsz_pop2024 %>%
  filter(REGION_N == "CENTRAL REGION")
region_bbox <- st_bbox(region_selected)

stats <- boxplot.stats(region_selected$DEPENDENCY)
outlier_vals <- stats$out
outlier_sf <- region_selected[region_selected$DEPENDENCY %in% outlier_vals, ]

tmap_mode("view")

tm_shape(region_selected, bbox = region_bbox) +
  tm_fill("DEPENDENCY", id = "SUBZONE_N", popup.vars = c("Name" = "SUBZONE_N", "Dependency" = "DEPENDENCY")) +
  tm_borders() +
  tm_shape(outlier_sf) +
  tm_borders(col = "red", lwd = 2) +
  tm_view(set_zoom_limits = c(12,14))

tmap_mode("plot")
```


