---
title: "Hands-on Exercise 2"
author: "Jin Jiaying"
date: "03 Sep 2026"
date-modified: last-modified
date-format: "DD MMM YYYY"
---

# Loading the necessary R packages

```{r}
pacman::p_load(sf, terra, spatstat, tmap, rvest, tidyverse)
```

# **2.1 1st Order Spatial Point Patterns Analysis Methods**

## 2.1.1 Importing data

```{r}
mpsz_sf <- st_read("hands-on_ex2_data/MasterPlan2019SubzoneBoundaryNoSeaGEOJSON.geojson") %>% 
  st_zm(drop = TRUE, what = "ZM") %>% 
  st_transform(crs = 3414)

childcare_sf <- st_read("hands-on_ex2_data/ChildCareServices.geojson") %>% 
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 3414)
```

### 2.2.1.1 Wrangling Geospatial Data

```{r}
mpsz_cl <- mpsz_sf %>%
  filter(SUBZONE_N != "SOUTHERN GROUP",
         PLN_AREA_N != "WESTERN ISLANDS",
         PLN_AREA_N != "NORTH-EASTERN ISLANDS")

write_rds(mpsz_cl, 
          "hands-on_ex2_data/mpsz_cl.rds")
```

::: {.callout-note title="Note on Geojson vs KML file"}
A GeoJSON file is used instead of the KML file provided in the original exercise. Since the required attributes such as `REGION_N`, `PLN_AREA_N`, `SUBZONE_N`, and `SUBZONE_C` are stored within the `Description` field in the GeoJSON file, the KML-specific pre-processing step for extracting attributes from `Description` is omitted.

``` r
extract_kml_field <- function(html_text, field_name) {
  if (is.na(html_text) || html_text == "") return(NA_character_)
  
  page <- read_html(html_text)
  rows <- page %>% html_elements("tr")
  
  value <- rows %>%
    keep(~ html_text2(html_element(.x, "th")) == field_name) %>%
    html_element("td") %>%
    html_text2()
  
  if (length(value) == 0) NA_character_ else value
}

mpsz_sf <- mpsz_sf %>%
  mutate(
    REGION_N = map_chr(Description, extract_kml_field, "REGION_N"),
    PLN_AREA_N = map_chr(Description, extract_kml_field, "PLN_AREA_N"),
    SUBZONE_N = map_chr(Description, extract_kml_field, "SUBZONE_N"),
    SUBZONE_C = map_chr(Description, extract_kml_field, "SUBZONE_C")
  ) %>%
  select(-Name, -Description) %>%
  relocate(geometry, .after = last_col())
```
:::

#### Mapping the geospatial data sets

```{r}
plot(st_geometry(mpsz_cl), col = "lightgrey")
plot(st_geometry(childcare_sf), add = TRUE, pch = 20, cex = 0.7)
```

```{r}
tmap_mode('view')
tm_shape(childcare_sf) + tm_dots()

tmap_mode('plot')
```

::: {.callout-note title="Note on Interactive Mapping with tmap"}
In interactive mode, `tmap` uses the Leaflet for R framework to create interactive maps. This allows users to zoom, pan, and click individual spatial features to view their attribute information.

The basemap can also be changed to different map providers, such as `ESRI.WorldGrayCanvas`, `OpenStreetMap`, and `ESRI.WorldTopoMap`.

After exploring the interactive map, it is recommended to switch `tmap` back to plot mode:

``` r
tmap_mode("plot")
```

Interactive maps require additional connections and resources, so excessive use of interactive maps should be avoided when rendering or publishing the Quarto website.
:::

## 2.1.2 Geospatial Data wrangling

```{r}
childcare_ppp <- as.ppp(childcare_sf)

class(childcare_ppp)
```

```{r}
summary(childcare_ppp)
```

```{r}
sg_owin <- as.owin(mpsz_cl)

class(sg_owin)

plot(sg_owin)
```

```{r}
childcareSG_ppp = childcare_ppp[sg_owin]

childcareSG_ppp
```

## 2.1.3 Clark-Evan Test for Nearest Neighbour Analysis

```{r}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion= sg_owin,
                alternative=c("clustered"))
```

::: {.callout-tip title="Interpretation of Clark-Evans test without CSR"}
**Statistical Conclusion**

The Clark-Evans test returns an R value of 0.53532, which is below 1, indicating a clustered spatial pattern. The p-value is less than 2.2e-16, which is substantially smaller than the 0.05 significance level. Therefore, the null hypothesis of Complete Spatial Randomness (CSR) is rejected.

There is statistically significant evidence that childcare centres in Singapore are spatially clustered rather than randomly distributed.

**Business Communication**

From a planning perspective, childcare centres tend to concentrate in particular parts of Singapore rather than being evenly distributed across the country. This clustering may reflect the concentration of residential areas and demand for childcare services. However, further analysis using demographic information, such as the distribution of young children and residential population, would be required to determine whether some areas are underserved.
:::

```{r}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion= sg_owin,
                alternative=c("clustered"),
                method="MonteCarlo",
                nsim=99)
```

::: {.callout-tip title="Interpretation of Clark-Evans test with CSR"}
**Statistical Conclusion**

The Clark-Evans test returns an R value of 0.53532, which is below 1, indicating a clustered spatial pattern. Using the Monte Carlo method with 99 simulations under Complete Spatial Randomness (CSR), the test produces a p-value of 0.01, which is smaller than the 0.05 significance level. Therefore, the null hypothesis of Complete Spatial Randomness is rejected.

There is statistically significant evidence that childcare centres in Singapore are spatially clustered. The Monte Carlo simulation further supports the conclusion that the observed clustering is unlikely to have occurred simply by chance under a random spatial distribution.

**Business Communication**

From a planning perspective, childcare centres tend to be concentrated in particular parts of Singapore rather than being randomly distributed across the country. The Monte Carlo test suggests that this clustering pattern is unlikely to be explained by random variation alone. This concentration may reflect differences in residential density, population distribution, and demand for childcare services across Singapore. However, additional demographic information, such as the number and distribution of young children, should be incorporated before determining whether particular areas are underserved.
:::

## 2.1.4 Kernel Density Estimation Method

```{r}
kde_SG_diggle <- density(
  childcareSG_ppp,
  sigma=bw.diggle,
  edge=TRUE,
  kernel="gaussian") 

plot(kde_SG_diggle)
```

```{r}
summary(kde_SG_diggle)
```

```{r}
bw <- bw.diggle(childcareSG_ppp)
bw
```

```{r}
childcareSG_ppp_km <- rescale.ppp(childcareSG_ppp, 1000, "km")

kde_childcareSG_km <- density(childcareSG_ppp_km,
                              sigma=bw.diggle,
                              edge=TRUE,
                              kernel="gaussian")

plot(kde_childcareSG_km)
```

```{r}
bw.CvL(childcareSG_ppp_km)

bw.scott(childcareSG_ppp_km)

bw.ppl(childcareSG_ppp_km)

bw.diggle(childcareSG_ppp_km)

kde_childcareSG.ppl <- density(childcareSG_ppp_km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG_km, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```

```{r}
par(mfrow=c(2,2))

plot(density(childcareSG_ppp_km, 
             sigma=0.2959712, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")

plot(density(childcareSG_ppp_km, 
             sigma=0.2959712, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")

plot(density(childcareSG_ppp_km, 
             sigma=0.2959712, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")

plot(density(childcareSG_ppp_km, 
             sigma=0.2959712, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```

## 2.1.5 Fixed and Adaptive KDE

```{r}
kde_childcareSG_fb <- density(childcareSG_ppp_km,
                              sigma=0.6, 
                              edge=TRUE,
                              kernel="gaussian")
plot(kde_childcareSG_fb)
```

```{r}
kde_childcareSG_ab <- adaptive.density(
  childcareSG_ppp_km, 
  method="kernel")
plot(kde_childcareSG_ab)
```

```{r}
par(mfrow=c(1,2))
plot(kde_childcareSG_fb, main = "Fixed bandwidth")
plot(kde_childcareSG_ab, main = "Adaptive bandwidth")
```

## 2.1.6 Plotting cartographic quality KDE map

```{r}
kde_childcareSG_bw_terra <- rast(kde_childcareSG_km)

class(kde_childcareSG_bw_terra)

kde_childcareSG_bw_terra

crs(kde_childcareSG_bw_terra) <- "EPSG:3414"

kde_childcareSG_bw_terra
```

```{r}
tm_shape(kde_childcareSG_bw_terra) + 
  tm_raster(col.scale = 
              tm_scale_continuous(
                values = "viridis"),
            col.legend = tm_legend(
            title = "Density values",
            title.size = 0.7,
            text.size = 0.7,
            bg.color = "white",
            bg.alpha = 0.7,
            position = tm_pos_in(
              "right", "bottom"),
            frame = TRUE)) +
  tm_graticules(labels.size = 0.7) + tm_compass() + tm_layout(scale = 1.0)
```

## 2.1.7 First Order SPPA at the Planning Subzone Level

```{r}
pg <- mpsz_cl %>%
  filter(PLN_AREA_N == "PUNGGOL")
tm <- mpsz_cl %>%
  filter(PLN_AREA_N == "TAMPINES")
ck <- mpsz_cl %>%
  filter(PLN_AREA_N == "CHOA CHU KANG")
jw <- mpsz_cl %>%
  filter(PLN_AREA_N == "JURONG WEST")

par(mfrow=c(2,2))
plot(st_geometry(pg), main = "Ponggol")
plot(st_geometry(tm), main = "Tampines")
plot(st_geometry(ck), main = "Choa Chu Kang")
plot(st_geometry(jw), main = "Jurong West")
```

```{r}
pg_owin = as.owin(pg)
tm_owin = as.owin(tm)
ck_owin = as.owin(ck)
jw_owin = as.owin(jw)

childcare_pg_ppp = childcare_ppp[pg_owin]
childcare_tm_ppp = childcare_ppp[tm_owin]
childcare_ck_ppp = childcare_ppp[ck_owin]
childcare_jw_ppp = childcare_ppp[jw_owin]

childcare_pg_ppp.km = rescale.ppp(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale.ppp(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale.ppp(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale.ppp(childcare_jw_ppp, 1000, "km")

par(mfrow=c(2,2))
plot(unmark(childcare_pg_ppp.km), 
  main="Punggol")
plot(unmark(childcare_tm_ppp.km), 
  main="Tampines")
plot(unmark(childcare_ck_ppp.km), 
  main="Choa Chu Kang")
plot(unmark(childcare_jw_ppp.km), 
  main="Jurong West")
```

```{r}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)

clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

```{r}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Jurong West")
```

# **2.2 2nd Order Spatial Point Patterns Analysis Methods**

## 2.2.1 Importing data

Same data be used as in 2.1.1

## 2.2.2 Analysing Spatial Point Process Using G-Function

```{r}
set.seed(1234)

G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```

```{r}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)

plot(G_CK.csr)
```

```{r}
G_tm = Gest(childcare_tm_ppp, correction = "best")
plot(G_tm)
```

```{r}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)
plot(G_tm.csr)
```

## 2.2.3 Analysing Spatial Point Process Using F-Function

```{r}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```

```{r}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)
plot(F_CK.csr)
```

```{r}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```

```{r}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)
plot(F_tm.csr)
```

## 2.2.4 Analysing Spatial Point Process Using K-Function

```{r}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

```{r}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, local=TRUE)
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

```{r}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

```{r}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, local=TRUE)
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))
```

## 2.2.5 Analysing Spatial Point Process Using L-Function

```{r}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

```{r}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, local=TRUE)
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

```{r}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

```{r}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, local=TRUE)
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```

::: {.callout-note title="Comparison of Second-order Spatial Point Pattern Functions"}
**G-function**

- Main Focus: Event-to-event

- What It Measures: Distance from each event to its nearest neighbouring event

- Indication of Clustering: Observed G is **higher** than CSR

**F-function**

- Main Focus: Empty space

- What It Measures: Distance from a random location to the nearest event

- Indication of Clustering: Observed F is **lower** than CSR

**K-function**

- Main Focus: Neighbourhood intensity

- What It Measures: Number of neighbouring events within distance (r)

- Indication of Clustering: Observed K is **higher** than CSR

**L-function**

- Main Focus: Standardised K-function

- What It Measures: Transformed version of K for easier interpretation

- Indication of Clustering: (L(r)-r \> 0)

For a **clustered pattern**:

- **G**: points are closer to one another.

- **F**: clustering creates larger empty spaces between clusters.

- **K**: more neighbouring points are found within distance (r).

- **L**: a positive (L(r)-r) indicates clustering.

If the observed curve stays close to the CSR expectation or within the simulation envelope, the pattern is approximately random at that distance scale.
:::
