---
title: "Hands-on Exercise 3"
author: "Jin Jiaying"
date: "05 Sep 2026"
date-modified: last-modified
---

# Loading the necessary R packages
```{r}
pacman::p_load(sf, spatstat, sparr, tmap, stpp, tidyverse)
```

# 3 Spatio-Temporal Point Patterns Analysis

## 3.1 Importing and Preparing Study Area
```{r}
raw <- st_read(dsn="hands-on_ex3_data/raw_data", layer = "Batas_Wilayah_KelurahanDesa_10K_AR") %>%
  st_transform(crs = 32748)
names(raw)
head(raw)
```
```{r}
unique(
  raw$WADMPR[
    grepl("Bangka", raw$WADMPR, ignore.case = TRUE)
  ]
)
```
```{r}
kbb <- raw %>%
  filter(WADMPR == "Kepulauan Bangka Belitung")
nrow(kbb)
unique(kbb$WADMPR)
kbb %>%
  st_drop_geometry() %>%
  count(WADMKK, sort = TRUE)
```
::: {.callout-note title="Data Preparation"}
The sub-district boundary of Kepulauan Bangka Belitung is getting by filtering latest whole indonesia shapefile (raw). After checking number of Kabupaten/Kota, it matches with latest official data. In this case, although the number of region is 393 which is higher than the hands-on material given, the following exercise will continue by using this 393 regions data.
:::

```{r}
kbb_sf <- kbb %>%
  st_union() %>%
  st_zm(drop = TRUE, what = "ZM") %>%
  st_transform(crs = 32748)
```


```{r}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```
```{r}
class(kbb_owin)
```

## 3.2 Importing and Preparing Forest Fire data
```{r}
fire_raw <- read_csv("hands-on_ex3_data/modis_2024_Indonesia.csv")
fire_sf <- fire_raw %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = 32748)
```
::: {.callout-note title="Data used"}
latest 2024 data is used in this exercise.
:::

```{r}
fire_sf <- fire_sf %>% 
  mutate(DayofYear = yday(acq_date)) %>%
  mutate(Month_num = month(acq_date)) %>%
  mutate(Month_fac = month(acq_date, 
                           label = TRUE, 
                           abbr = FALSE))
```

## 3.3 Visualising the Fire Points
```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
tm_shape(fire_sf) +
  tm_dots() +
  tm_title("Overall plot") + 
  tm_credits("Forest Fires in Kepulauan Bangka Belitung")
```
::: {.callout-note title="Plot Title"}
tm_credit() is initially used as credit introduction, since tmap does not have function like sub-title, here it is used just for web display.
:::

```{r}
tm_shape(kbb_sf)+
  tm_polygons() +
tm_shape(fire_sf) +
  tm_dots(size = 0.1) +
tm_facets(by="Month_fac", 
            free.coords=FALSE, 
            drop.units = TRUE) + 
  tm_title("Monthly geographic distribution of forest fires")
```

## 3.4 Computing STKDE by Month
```{r}
fire_month <- fire_sf %>% 
  select(Month_num)

fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp

summary(fire_month_ppp)

any(duplicated(fire_month_ppp))
```
```{r}
fire_month_owin <- fire_month_ppp[kbb_owin]
summary(fire_month_owin)

plot(fire_month_owin)
```
```{r}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```
```{r}
tims <- c(7,8,9,10,11,12)
par(mfcol=c(2,3))
for(i in tims){ 
  plot(st_kde, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at month",i))
}
```

## 3.5 Computing STKDE by Day of Year
```{r}
fire_yday_ppp <- fire_sf %>% 
  select(DayofYear) %>%
  as.ppp()

fire_yday_owin <- fire_yday_ppp[kbb_owin]
summary(fire_yday_owin)

kde_yday <- spattemp.density(
  fire_yday_owin)
summary(kde_yday)

plot(kde_yday)
```

## 3.6 Computing STKDE by Day of Year: Improved method
```{r}
set.seed(1234)
BOOT.spattemp(fire_yday_owin)
```
```{r}
kde_yday_improved <- spattemp.density(fire_yday_owin,
                             h = 9000,
                             lambda = 19)
summary(kde_yday_improved)
```
```{r}
plot(kde_yday_improved)
```

## 3.7 Spatio-temporal Point Patterns Analysis: stpp methods
```{r}
coords <- st_coordinates(fire_sf)

fire_df <- data.frame(
  x = coords[, 1],  
  y = coords[, 2],
  t = fire_sf$`DayofYear`)

fire_stpp <- as.3dpoints(fire_df)
class(fire_stpp)

plot(fire_stpp)
```

```{r}
kbb_stik <- STIKhat(fire_stpp)

plotK(kbb_stik)
```
::: {.callout-note title="Warning on duplicated spatial points"}
STIKhat() returns a warning indicating duplicated spatial points. 
A check of the input data found no duplicated space-time observations 
(x, y, t), while 12 observations share spatial coordinates (x, y) with 
another observation but occur at different times. Therefore, these 
observations are retained for the space-time analysis.
:::
