---
title: "Take-home Exercise 1 - Technical Report"
author: "Jin Jiaying"
date: "05 Sep 2026"
date-modified: last-modified
---

# Overview

This is Technical Report of Take-home Exercise 1, aiming to examine the spatial patterns of road traffic accidents in the Bangkok Greater Metropolitan Area (BGMA) in 2022. Dataset from Kaggle [Thailand Road Accident 2019-2022](https://www.kaggle.com/datasets/thaweewatboy/thailand-road-accident-2019-2022) is used. Administrative boundary data and road network data will also be incorporated to support spatial and network-constrained analyses.

The analysis consists of four main tasks:

- Data Preparation and Quality Assessment: Clean and validate the accident records, define the study area, handle missing or invalid coordinates, and prepare the spatial datasets for analysis.

- Exploratory and First-Order Analysis: Examine the spatial and temporal distribution of accidents and identify areas or road segments with high accident intensity using methods such as Kernel Density Estimation (KDE) and Network Kernel Density Estimation (NKDE).

- Second-Order Analysis – Investigate whether road accidents exhibit statistically significant spatial clustering, with particular attention to clustering along the road network.

- Integration and Interpretation – Compare findings from conventional and network-constrained approaches and discuss their implications for road-safety planning and public-safety management.


# Loading Necessary R Packages
```{r}
pacman::p_load(tidyverse, sf, lubridate, spNetwork)
```


# 1 Data Preparation

## 1.1 Data Import and Initial Quality Assessment

### 1.1.1 Data Import
```{r}
th_accident_raw <- read_csv("take-home_ex1_data/thai_road_accident_2019_2022.csv")
```

### 1.1.2 Restrict to 2022
```{r}
accident_2022 <- th_accident_raw %>%
  filter(year(incident_datetime) == 2022)
```

### 1.1.3 Initial Data Screening
```{r}
glimpse(accident_2022)
summary(accident_2022)
```
:::
:::

### 1.1.4 Missing Value Assessment
```{r}
colSums(is.na(accident_2022))
```

### 1.1.5 Duplicate Record Assessment
```{r}
sum(duplicated(accident_2022))

accident_2022 %>%
  count(acc_code) %>%
  filter(n > 1)
```
No duplicated accident identifiers or fully duplicated records were detected.


## 1.2 Study Area Definition

Bangkok Province was selected as the study area because it provides a clearly defined urban road-safety case while retaining the complete set of accident events within the selected geographic extent. Restricting the analysis to Bangkok also provides a computationally tractable road network for network-constrained spatial point-pattern analysis without requiring random subsampling of accident events.

### 1.2.1 Select Bangkok Accident Records
```{r}
accident_bkk <- accident_2022 %>%
  filter(province_en == "Bangkok")
```

### 1.2.2 Prepare Bangkok Administrative Boundary
```{r}
tha_adm1 <- st_read("take-home_ex1_data/geoBoundaries-THA-ADM1.geojson")

bkk_boundary <- tha_adm1 %>%
  filter(shapeName == "Bangkok")
```
:::
:::


## 1.3 Event and Attribute Preparation

### 1.3.1 Define accident events
A point event represents one recorded road traffic accident occurring within the Bangkok Metropolitan Region during 2022. Each unique acc_code is treated as one event. Multiple accidents occurring at the same recorded coordinates are retained as separate events because they represent distinct accident incidents.

### 1.3.2 Derive Temporal Attributes
```{r}
accident_bkk <- accident_bkk %>%
  mutate(
    month = month(incident_datetime, label = TRUE),
    weekday = wday(incident_datetime, label = TRUE),
    hour = hour(incident_datetime))
```

### 1.3.3 Derive Accident Severity
```{r}
accident_bkk <- accident_bkk %>%
  mutate(
    severity = case_when(
      is.na(number_of_fatalities) | is.na(number_of_injuries) ~ "Unknown",
      number_of_fatalities > 0 ~ "Fatal",
      number_of_injuries > 0 ~ "Injury",
      TRUE ~ "Non-injury"))
```


## 1.4 Geospatial Data Preparation

### 1.4.1 Coordinate Completeness and Plausibility
```{r}
missing_coord <- accident_bkk %>%
  filter(is.na(latitude) | is.na(longitude))

nrow(missing_coord)
```
:::
Records with missing coordinates were retained during the initial quality assessment but excluded from spatial point-pattern analysis because their locations could not be reliably represented spatially.
:::

```{r}
accident_bkk_clean <- accident_bkk %>%
  filter(
    !is.na(latitude),
    !is.na(longitude))
```

### 1.4.2 Convert Records to Spatial Point Features
```{r}
accident_bkk_sf <- st_as_sf(
  accident_bkk_clean,
  coords = c("longitude", "latitude"),
  crs = 4326,
  remove = FALSE)
```

### 1.4.3 Spatial Validation Against bkk Boundary
```{r}
accident_bkk_sf <- st_filter(
  accident_bkk_sf,
  bkk_boundary)
```

### 1.4.4 CRS Transformation
```{r}
accident_bkk_utm <- st_transform(
  accident_bkk_sf,
  32647)

bkk_boundary_utm <- st_transform(
  bkk_boundary,
  32647)
```


## 1.5 Road Network Preparation
The road network represents an OpenStreetMap snapshot near the end of the 2022 study period; therefore, minor changes to the network during 2022 may not be fully represented.(https://download.geofabrik.de/asia/thailand.html)

### 1.5.1 Import and Inspect Road Network
```{r}
roads_th <- st_read("take-home_ex1_data/th_road/gis_osm_roads_free_1.shp")

roads_th
st_crs(roads_th)
unique(st_geometry_type(roads_th))
unique(roads_th$fclass)
```

### 1.5.2 Select Relevant Road Classes
```{r}
motor_road_classes <- c(
  "motorway",
  "motorway_link",
  "trunk",
  "trunk_link",
  "primary",
  "primary_link",
  "secondary",
  "secondary_link",
  "tertiary",
  "tertiary_link",
  "unclassified",
  "residential",
  "living_street",
  "service")

roads_motor <- roads_th %>%
  filter(fclass %in% motor_road_classes)

roads_motor %>%
  st_drop_geometry() %>%
  count(fclass, sort = TRUE)
```

### 1.5.3 Restrict Road Network to BKK
```{r}
st_crs(roads_motor)
st_crs(bkk_boundary)

st_crs(roads_motor) == st_crs(bkk_boundary)
```

```{r}
roads_bkk <- st_filter(
  roads_motor,
  bkk_boundary)

nrow(roads_bkk)
roads_bkk %>%
  st_drop_geometry() %>%
   count(fclass, sort = TRUE)
```

### 1.5.4 Clip Network to bkk Boundary
```{r}
roads_bkk <- st_intersection(
  roads_bkk,
  bkk_boundary)
```

### 1.5.5 Network Geometry Preparation and Validation
```{r}
roads_bkk <- roads_bkk %>%
  st_cast("LINESTRING")
unique(st_geometry_type(roads_bkk))
```

```{r}
table(st_is_valid(roads_bkk))
nrow(roads_bkk)
```

```{r}
roads_bkk <- roads_bkk %>%
  mutate(LineID = row_number())

anyDuplicated(roads_bkk$LineID)
```

```{r}
roads_bkk_utm <- st_transform(
  roads_bkk,
  32647)
```


## 1.6 Point-to-Network Validation and Snapping

### 1.6.1 Point-to-Network Distance Assessment
```{r}
nearest_road_id <- st_nearest_feature(
  accident_bkk_utm,
  roads_bkk_utm)

nearest_road_dist <- st_distance(
  accident_bkk_utm,
  roads_bkk_utm[nearest_road_id, ],
  by_element = TRUE)
```

```{r}
accident_bkk_utm$dist_to_road_m <-
  as.numeric(nearest_road_dist)

summary(accident_bkk_utm$dist_to_road_m)
quantile(
  accident_bkk_utm$dist_to_road_m,
  probs = c(
    0,
    0.25,
    0.5,
    0.75,
    0.90,
    0.95,
    0.99,
    1),
  na.rm = TRUE)

accident_bkk_utm %>%
  st_drop_geometry() %>%
  summarise(
    total = n(),
    over_10m = sum(dist_to_road_m > 10),
    over_20m = sum(dist_to_road_m > 20),
    over_30m = sum(dist_to_road_m > 30),
    over_50m = sum(dist_to_road_m > 50))
```
:::
Point-to-Network Validation
Before conducting network-constrained spatial analysis, the distance from each accident point to its nearest road segment was calculated to assess positional consistency. The median distance was approximately 
:::

### 1.6.2 Examine Potentially Misaligned Events
```{r}
far_accidents <- accident_bkk_utm %>%
  filter(dist_to_road_m > 20)

far_accidents %>%
  st_drop_geometry() %>%
  select(
    acc_code,
    latitude,
    longitude,
    dist_to_road_m,
    route,
    road_description) %>%
  arrange(desc(dist_to_road_m))
```

```{r}
far_nearest_id <- st_nearest_feature(
  far_accidents,
  roads_bkk_utm)

nearest_lines <- st_nearest_points(
  far_accidents,
  roads_bkk_utm[far_nearest_id, ],
  pairwise = TRUE)

far_area <- st_bbox(
  st_buffer(
    st_union(far_accidents),
    dist = 500))

plot(
  st_geometry(roads_bkk_utm),
  xlim = c(far_area["xmin"], far_area["xmax"]),
  ylim = c(far_area["ymin"], far_area["ymax"]),
  col = "grey")

plot(
  nearest_lines,
  add = TRUE,
  lwd = 2)

plot(
  st_geometry(far_accidents),
  add = TRUE,
  pch = 19)
```
:::
Accident events located more than 20 m from their nearest road segment were visually inspected to identify potential spatial misalignment before network snapping.
:::

### 1.6.3 Snap Accident Events to Road Network
```{r}
accident_snapped <- snapPointsToLines2(
  points = accident_bkk_utm,
  lines = roads_bkk_utm,
  idField = "LineID",
  snap_dist = 30)
```

### 1.6.4 Post-Snapping Validation
```{r}
nrow(accident_bkk_utm)
nrow(accident_snapped)

unique(st_geometry_type(accident_snapped))

st_crs(accident_snapped) == st_crs(roads_bkk_utm)

sum(is.na(accident_snapped$nearest_line_id))

accident_snapped %>%
  st_drop_geometry() %>%
  select(acc_code, nearest_line_id) %>%
  head()
```

```{r}
check_road_id <- st_nearest_feature(
  accident_snapped,
  roads_bkk_utm)

check_distance <- st_distance(
  accident_snapped,
  roads_bkk_utm[check_road_id, ],
  by_element = TRUE)

summary(as.numeric(check_distance))
```


## 1.7 Store Analysis-Ready Data
```{r}
analysis_ready <- list(
  accidents = accident_snapped,
  roads = roads_bkk_utm,
  boundary = bkk_boundary_utm)

saveRDS(
  analysis_ready,
  "take-home_ex1_data/processed/analysis_ready.rds")
```


# 2 Network-Constrained Spatial Point-Pattern Analysis (SPPA)

```{r}
analysis_ready <- readRDS("take-home_ex1_data/processed/analysis_ready.rds")

accident_snapped <- analysis_ready$accidents
roads_bkk_utm <- analysis_ready$roads
bkk_boundary_utm <- analysis_ready$boundary
```

## 2.1 First-Order Network-Constrained SPPA (NKDE)

### 2.1.1 Preparing Lixels and Sampling Points
```{r}
road_length_m <- as.numeric(
  st_length(roads_bkk_utm))

summary(road_length_m)

quantile(
  road_length_m,
  probs = c(
    0,
    0.25,
    0.50,
    0.75,
    0.90,
    0.95,
    0.99,
    1),
  na.rm = TRUE)
```
:::
Lixel length selection. The distribution of the original road-segment lengths was examined before lixelisation. The median road-segment length was approximately 65 m, while the 90th and 95th percentiles were approximately 306 m and 487 m, respectively. A lixel length of 300 m was therefore adopted as an initial sampling resolution. This subdivides relatively long road segments while avoiding unnecessary fragmentation of the already highly segmented urban road network.

A minimum residual lixel length of 50 m was specified to prevent very short remainder segments from being created during the subdivision process.
:::

```{r}
lixel_length <- 300
min_lixel_length <- 50

lixels <- lixelize_lines(
  lines = roads_bkk_utm,
  lx_length = lixel_length,
  mindist = min_lixel_length)
```

```{r}
nrow(roads_bkk_utm)
nrow(lixels)

lixel_length_m <- as.numeric(st_length(lixels))

summary(lixel_length_m)
range(lixel_length_m)
```
:::
ixelisation. A lixel length of 300 m with a minimum residual length of 50 m was applied. The number of network segments increased moderately from 208,445 to 231,380 after lixelisation, indicating that the selected resolution did not cause excessive fragmentation. The resulting lixels had a median length of approximately 75 m and a maximum length of approximately 350 m. The latter is consistent with the minimum-distance rule, whereby a short residual segment may be merged with the preceding lixel.
:::

```{r}
samples <- lines_center(lixels)
```

### 2.1.2 Bandwidth Selection / Sensitivity Analysis
```{r}
run_nkde <- function(bw_value) {
  
  nkde(
    lines = roads_bkk_utm,
    events = accident_bkk_utm,
    w = rep(1, nrow(accident_bkk_utm)),
    samples = samples,
    
    kernel_name = "quartic",
    bw = bw_value,
    adaptive = FALSE,
    
    method = "simple",
    div = "bw",
    
    max_depth = 8,
    digits = 1,
    tol = 1,
    
    agg = NULL,
    sparse = TRUE,
    grid_shape = c(10, 10),
    
    verbose = TRUE,
    check = TRUE
  )
}
```

```{r}
nkde_500 <- run_nkde(500)
```

```{r}
length(nkde_500) == nrow(samples)
```

```{r}
nkde_1000 <- run_nkde(1000)
```
```{r}
length(nkde_1000) == nrow(samples)
```

```{r}
nkde_compare <- tibble(
  lixel_id = seq_len(nrow(samples)),
  bw_500 = as.numeric(nkde_500) * 1000,
  bw_1000 = as.numeric(nkde_1000) * 1000
)

nkde_compare %>%
  summarise(
    across(
      c(bw_500, bw_1000),
      list(
        median = ~ median(.x, na.rm = TRUE),
        p90 = ~ quantile(.x, 0.90, na.rm = TRUE),
        p95 = ~ quantile(.x, 0.95, na.rm = TRUE),
        p99 = ~ quantile(.x, 0.99, na.rm = TRUE),
        max = ~ max(.x, na.rm = TRUE)
      )
    )
  )
```

```{r}
cut_500 <- quantile(
  nkde_compare$bw_500,
  0.95,
  na.rm = TRUE
)

cut_1000 <- quantile(
  nkde_compare$bw_1000,
  0.95,
  na.rm = TRUE
)

nkde_compare <- nkde_compare %>%
  mutate(
    high_500 = bw_500 >= cut_500,
    high_1000 = bw_1000 >= cut_1000
  )


table(
  nkde_compare$high_500,
  nkde_compare$high_1000
)

intersection_n <- sum(
  nkde_compare$high_500 &
  nkde_compare$high_1000
)

union_n <- sum(
  nkde_compare$high_500 |
  nkde_compare$high_1000
)

hotspot_overlap <- intersection_n / union_n

hotspot_overlap
```
The high-intensity road segments identified using 500 m and 1,000 m bandwidths showed substantial spatial agreement. Approximately 78% of the lixels classified within the top 5% of network intensity were shared between the two bandwidths, corresponding to a Jaccard overlap of approximately 0.63. This suggests that the identification of the principal high-intensity locations was reasonably robust to the selected bandwidth, although some local differences remained.

#### Bandwidth Sensitivity Analysis
```{r}
lixels_compare <- lixels %>%
  mutate(
    bw_500 = nkde_compare$bw_500,
    bw_1000 = nkde_compare$bw_1000
  )

common_limit <- quantile(
  c(
    lixels_compare$bw_500,
    lixels_compare$bw_1000
  ),
  0.99,
  na.rm = TRUE
)

#500
ggplot() +
  geom_sf(
    data = lixels_compare,
    aes(colour = bw_500),
    linewidth = 0.4
  ) +
  geom_sf(
    data = bkk_boundary_utm,
    fill = NA,
    colour = "grey30",
    linewidth = 0.3
  ) +
  scale_colour_viridis_c(
    limits = c(0, common_limit),
    oob = scales::squish,
    name = "Accident intensity\n(per km)"
  ) +
  theme_void() +
  labs(
    title = "NKDE of Road Accidents",
    subtitle = "Bangkok, 2022 | Bandwidth = 500 m"
  )

#1000
ggplot() +
  geom_sf(
    data = lixels_compare,
    aes(colour = bw_1000),
    linewidth = 0.4
  ) +
  geom_sf(
    data = bkk_boundary_utm,
    fill = NA,
    colour = "grey30",
    linewidth = 0.3
  ) +
  scale_colour_viridis_c(
    limits = c(0, common_limit),
    oob = scales::squish,
    name = "Accident intensity\n(per km)"
  ) +
  theme_void() +
  labs(
    title = "NKDE of Road Accidents",
    subtitle = "Bangkok, 2022 | Bandwidth = 1000 m"
  )

```
:::
NKDE results using bandwidths of 500 m and 1,000 m were compared to assess the sensitivity of the estimated accident-intensity pattern to the degree of smoothing. The two bandwidths identified broadly similar high-intensity locations, with approximately 78% of the top 5% high-intensity lixels shared between the two results. The corresponding Jaccard overlap was approximately 0.63, indicating substantial spatial agreement. Visually, the 500 m bandwidth retained greater local variation, whereas the 1,000 m bandwidth produced more continuous road-corridor patterns without substantially altering the principal high-intensity locations. The 1,000 m bandwidth was therefore adopted for the subsequent analysis.
:::

```{r}
final_nkde <- nkde_1000

lixels_nkde <- lixels %>%
  mutate(
    nkde = as.numeric(final_nkde),
    nkde_km = nkde * 1000
  )
```


### 2.1.3 Network Kernel Density Estimation
done

### 2.1.4 Spatial Distribution of Network Accident Intensity
```{r}
summary(lixels_nkde$nkde_km)

quantile(
  lixels_nkde$nkde_km,
  probs = c(
    0,
    0.25,
    0.50,
    0.75,
    0.90,
    0.95,
    0.99,
    1
  ),
  na.rm = TRUE
)
```
:::
The NKDE intensity distribution was highly right-skewed. At least 75% of the lixels had an estimated intensity of zero, while the 95th and 99th percentiles increased to approximately 0.00095 and 0.00353 respectively, indicating that elevated accident intensity was concentrated within a relatively small proportion of the analysed road network.
:::

```{r}
# 95th percentile threshold
p95_nkde <- quantile(
  lixels_nkde$nkde_km,
  probs = 0.95,
  na.rm = TRUE
)

# Extract top 5% high-intensity lixels
high_intensity_lixels <- lixels_nkde %>%
  filter(nkde_km >= p95_nkde)
```


## 2.2 Second-Order Network-Constrained SPPA (Network G/K function)

### 2.2.1 Network G- and K-Function


### 2.2.2 Complete Spatial Randomness Test


### 2.2.3 Monte Carlo Simulation Envelope


### 2.2.4 Interpretation of Network-distance Scales



## 2.3 Integration and Interpretation



# 3 Geovisualisation and Analytical Interpretation


# 4 Public-Safety Planning and Management


