---
title: "Take-home Exercise 1 - Technical Report"
author: "Jin Jiaying"
date: "05 Sep 2026"
date-modified: last-modified
---

# Overview

This is Technical Report of Take-home Exercise 1, aiming to examine the spatial patterns of road traffic accidents in the Bangkok Greater Metropolitan Area (BGMA) in 2022. Dataset from Kaggle [Thailand Road Accident 2019-2022](https://www.kaggle.com/datasets/thaweewatboy/thailand-road-accident-2019-2022) is used. Administrative boundary data and road network data will also be incorporated to support spatial and network-constrained analyses.

The analysis consists of four main tasks:

- Data Preparation and Quality Assessment: Clean and validate the accident records, define the study area, handle missing or invalid coordinates, and prepare the spatial datasets for analysis.

- Exploratory and First-Order Analysis: Examine the spatial and temporal distribution of accidents and identify areas or road segments with high accident intensity using methods such as Kernel Density Estimation (KDE) and Network Kernel Density Estimation (NKDE).

- Second-Order Analysis – Investigate whether road accidents exhibit statistically significant spatial clustering, with particular attention to clustering along the road network.

- Integration and Interpretation – Compare findings from conventional and network-constrained approaches and discuss their implications for road-safety planning and public-safety management.


# Loading Necessary R Packages
```{r}
pacman::p_load(tidyverse, sf, lubridate, spNetwork)
```


# 1 Data Preparation

## 1.1 Data Import and Initial Quality Assessment

### 1.1.1 Data Import
```{r}
th_accident_raw <- read_csv("take-home_ex1_data/thai_road_accident_2019_2022.csv")
```

### 1.1.2 Restrict to 2022
```{r}
accident_2022 <- th_accident_raw %>%
  filter(year(incident_datetime) == 2022)
```

### 1.1.3 Initial Data Screening
```{r}
glimpse(accident_2022)
summary(accident_2022)
```
:::
:::

### 1.1.4 Missing Value Assessment
```{r}
colSums(is.na(accident_2022))
```

### 1.1.5 Duplicate Record Assessment
```{r}
sum(duplicated(accident_2022))

accident_2022 %>%
  count(acc_code) %>%
  filter(n > 1)
```
No duplicated accident identifiers or fully duplicated records were detected.


## 1.2 Study Area Definition

### 1.2.1 Define Bangkok Metropolitan Region (BMR)
[Insert source]
```{r}
bmr_provinces <- c(
  "Bangkok",
  "Nakhon Pathom",
  "Nonthaburi",
  "Pathum Thani",
  "Samut Prakan",
  "Samut Sakhon")
```

### 1.2.2 Select BMR Accident Records
```{r}
accident_bmr <- accident_2022 %>%
  filter(province_en %in% bmr_provinces)
```

### 1.2.3 Prepare BMR Administrative Boundary
```{r}
tha_adm1 <- st_read("take-home_ex1_data/geoBoundaries-THA-ADM1.geojson")

bmr_names <- c(
  "Bangkok",
  "Nakhon Pathom Province",
  "Nonthaburi Province",
  "Pathum Thani Province",
  "Samut Prakan Province",
  "Samut Sakhon Province")

bmr_adm1 <- tha_adm1 %>%
  filter(shapeName %in% bmr_names)
```

### 1.2.4 Construct BMR Study Boundary
```{r}
bmr_boundary <- bmr_adm1 %>%
  summarise()
```
:::
:::


## 1.3 Event and Attribute Preparation

### 1.3.1 Define accident events
A point event represents one recorded road traffic accident occurring within the Bangkok Metropolitan Region during 2022. Each unique acc_code is treated as one event. Multiple accidents occurring at the same recorded coordinates are retained as separate events because they represent distinct accident incidents.

### 1.3.2 Derive Temporal Attributes
```{r}
accident_bmr <- accident_bmr %>%
  mutate(
    month = month(incident_datetime, label = TRUE),
    weekday = wday(incident_datetime, label = TRUE),
    hour = hour(incident_datetime))
```

### 1.3.3 Derive Accident Severity
```{r}
accident_bmr <- accident_bmr %>%
  mutate(
    severity = case_when(
      is.na(number_of_fatalities) | is.na(number_of_injuries) ~ "Unknown",
      number_of_fatalities > 0 ~ "Fatal",
      number_of_injuries > 0 ~ "Injury",
      TRUE ~ "Non-injury"))
```


## 1.4 Geospatial Data Preparation

### 1.4.1 Coordinate Completeness and Plausibility
```{r}
missing_coord <- accident_bmr %>%
  filter(is.na(latitude) | is.na(longitude))

nrow(missing_coord)
```
:::
Records with missing coordinates were retained during the initial quality assessment but excluded from spatial point-pattern analysis because their locations could not be reliably represented spatially.
:::

```{r}
accident_bmr_clean <- accident_bmr %>%
  filter(
    !is.na(latitude),
    !is.na(longitude))
```

### 1.4.2 Convert Records to Spatial Point Features
```{r}
accident_bmr_sf <- st_as_sf(
  accident_bmr_clean,
  coords = c("longitude", "latitude"),
  crs = 4326,
  remove = FALSE)
```

### 1.4.3 Spatial Validation Against BMR Boundary
```{r}
accident_bmr_sf <- st_filter(
  accident_bmr_sf,
  bmr_boundary)
```

### 1.4.4 CRS Transformation
```{r}
accident_bmr_utm <- st_transform(
  accident_bmr_sf,
  32647)

bmr_boundary_utm <- st_transform(
  bmr_boundary,
  32647)
```


## 1.5 Road Network Preparation
The road network represents an OpenStreetMap snapshot near the end of the 2022 study period; therefore, minor changes to the network during 2022 may not be fully represented.(https://download.geofabrik.de/asia/thailand.html)

### 1.5.1 Import and Inspect Road Network
```{r}
roads_th <- st_read("take-home_ex1_data/th_road/gis_osm_roads_free_1.shp")

roads_th
st_crs(roads_th)
unique(st_geometry_type(roads_th))
unique(roads_th$fclass)
```

### 1.5.2 Select Relevant Road Classes
```{r}
motor_road_classes <- c(
  "motorway",
  "motorway_link",
  "trunk",
  "trunk_link",
  "primary",
  "primary_link",
  "secondary",
  "secondary_link",
  "tertiary",
  "tertiary_link",
  "unclassified",
  "residential",
  "living_street",
  "service")

roads_motor <- roads_th %>%
  filter(fclass %in% motor_road_classes)

roads_motor %>%
  st_drop_geometry() %>%
  count(fclass, sort = TRUE)
```

### 1.5.3 Restrict Road Network to BMR
```{r}
st_crs(roads_motor)
st_crs(bmr_boundary)

st_crs(roads_motor) == st_crs(bmr_boundary)
```

```{r}
roads_bmr <- st_filter(
  roads_motor,
  bmr_boundary)

nrow(roads_bmr)
roads_bmr %>%
  st_drop_geometry() %>%
   count(fclass, sort = TRUE)
```

### 1.5.4 Clip Network to BMR Boundary
```{r}
roads_bmr <- st_intersection(
  roads_bmr,
  bmr_boundary)
```

### 1.5.5 Network Geometry Preparation and Validation
```{r}
roads_bmr <- roads_bmr %>%
  st_cast("LINESTRING")
unique(st_geometry_type(roads_bmr))
```

```{r}
table(st_is_valid(roads_bmr))
nrow(roads_bmr)
```

```{r}
roads_bmr <- roads_bmr %>%
  mutate(LineID = row_number())

anyDuplicated(roads_bmr$LineID)
```

```{r}
roads_bmr_utm <- st_transform(
  roads_bmr,
  32647)
```


## 1.6 Point-to-Network Validation and Snapping

### 1.6.1 Point-to-Network Distance Assessment
```{r}
nearest_road_id <- st_nearest_feature(
  accident_bmr_utm,
  roads_bmr_utm)

nearest_road_dist <- st_distance(
  accident_bmr_utm,
  roads_bmr_utm[nearest_road_id, ],
  by_element = TRUE)
```

```{r}
accident_bmr_utm$dist_to_road_m <-
  as.numeric(nearest_road_dist)

summary(accident_bmr_utm$dist_to_road_m)
quantile(
  accident_bmr_utm$dist_to_road_m,
  probs = c(
    0,
    0.25,
    0.5,
    0.75,
    0.90,
    0.95,
    0.99,
    1),
  na.rm = TRUE)

accident_bmr_utm %>%
  st_drop_geometry() %>%
  summarise(
    total = n(),
    over_10m = sum(dist_to_road_m > 10),
    over_20m = sum(dist_to_road_m > 20),
    over_30m = sum(dist_to_road_m > 30),
    over_50m = sum(dist_to_road_m > 50))
```
:::
Point-to-Network Validation
Before conducting network-constrained spatial analysis, the distance from each accident point to its nearest road segment was calculated to assess positional consistency. The median distance was approximately 3.85 m, while 95% and 99% of accident locations were within approximately 10.58 m and 13.96 m of the nearest road, respectively. The maximum observed distance was approximately 41.53 m. These results suggest that the accident locations were generally well aligned with the road network. A maximum snapping tolerance of 50 m was therefore considered appropriate for projecting accident points onto the road network.
:::

### 1.6.2 Examine Potentially Misaligned Events
```{r}
far_accidents <- accident_bmr_utm %>%
  filter(dist_to_road_m > 20)

far_accidents %>%
  st_drop_geometry() %>%
  select(
    acc_code,
    latitude,
    longitude,
    dist_to_road_m,
    route,
    road_description) %>%
  arrange(desc(dist_to_road_m))
```

```{r}
far_nearest_id <- st_nearest_feature(
  far_accidents,
  roads_bmr_utm)

nearest_lines <- st_nearest_points(
  far_accidents,
  roads_bmr_utm[far_nearest_id, ],
  pairwise = TRUE)

far_area <- st_bbox(
  st_buffer(
    st_union(far_accidents),
    dist = 500))

plot(
  st_geometry(roads_bmr_utm),
  xlim = c(far_area["xmin"], far_area["xmax"]),
  ylim = c(far_area["ymin"], far_area["ymax"]),
  col = "grey")

plot(
  nearest_lines,
  add = TRUE,
  lwd = 2)

plot(
  st_geometry(far_accidents),
  add = TRUE,
  pch = 19)
```
:::
Accident events located more than 20 m from their nearest road segment were visually inspected to identify potential spatial misalignment before network snapping.
:::

### 1.6.3 Snap Accident Events to Road Network
```{r}
accident_snapped <- snapPointsToLines2(
  points = accident_bmr_utm,
  lines = roads_bmr_utm,
  idField = "LineID",
  snap_dist = 50)
```

### 1.6.4 Post-Snapping Validation
```{r}
nrow(accident_bmr_utm)
nrow(accident_snapped)

unique(st_geometry_type(accident_snapped))

st_crs(accident_snapped) == st_crs(roads_bmr_utm)

sum(is.na(accident_snapped$nearest_line_id))

accident_snapped %>%
  st_drop_geometry() %>%
  select(acc_code, nearest_line_id) %>%
  head()
```

```{r}
check_road_id <- st_nearest_feature(
  accident_snapped,
  roads_bmr_utm)

check_distance <- st_distance(
  accident_snapped,
  roads_bmr_utm[check_road_id, ],
  by_element = TRUE)

summary(as.numeric(check_distance))
```


## 1.7 Store Analysis-Ready Data
```{r}
analysis_ready <- list(
  accidents = accident_snapped,
  roads = roads_bmr_utm,
  boundary = bmr_boundary_utm)

saveRDS(
  analysis_ready,
  "take-home_ex1_data/processed/analysis_ready.rds")
```


# 2 First-Order and Second-Order Point-Pattern Analysis

## 2.1 First-Order Point-Pattern Analysis

## 2.2 Second-Order Point-Pattern Analysis

## 2.3 Integration and Interpretation


# 3 Geovisualisation and Analytical Interpretation


# 4 Public-Safety Planning and Management


